Four Models for Severe Weather Deaths

### 2017-05-11 presentation.tex
First Pass at ZIP Slide

	\section{Zero Inflated Poisson Model}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}

	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}

	\end{frame}

	\section{Zero Inflated Negative Binomial Model}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		The Zero Inflated Negative Binomial (ZINB) model has three parameters $\sigma,$ the probability of a structural zero, and $p,r$ the usual negative binomial parameters. 
		For a single $X$ the probability density is: 
		$$p(X|\sigma, p, r) = \sigma I_{X=0}(X) + (1-\sigma)\frac{\Gamma(r+X)}{\Gamma(r)X!}p^X(1-p)^r$$
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		We took the uniform priors on $\sigma$ and $p$ and the non-informative prior $r^{-1/2}$ on $r$.
		This gave a posterior of:
		\scriptsize$$p(r,\sigma, p|X)\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}\prod_{i=1}^N\left(\frac{\Gamma(r+X_i)}{r^{1/2}\Gamma(r)}\right)$$
		\normalsize This factors into nothing. Therefore we used the Metropolis Hastings Algorithm with proposal distributions of truncated normals.
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model:Results}
		\begin{itemize}
			\item Curse of Dimensionality; very slow convergence.
			\item High autocorrelation
			\item Low confidence in results
		\end{itemize}
		\begin{table}
			\centering
			\scalebox{0.6}{
			\begin{tabular}{lcccccc}
				\toprule
				&\multicolumn{3}{c}{Flash Flood}&\multicolumn{3}{c}{Tornado}\\
				\cmidrule(r){2-4}\cmidrule(l){5-7}
				Parameter & Mean Value &Median Value & (95\% CI)& Mean Value &Median Value & (95\% CI)\\
				\midrule
				$\sigma$ & 0.307 &
				0.282&
				(0.013, 
				0.716)&
				0.299 &
				0.276&
				(0.012, 
				0.702)\\
				$p$ &  0.554 &
				0.555&
				(0.518, 
				0.589)&
				0.862 &
				0.863&
				(0.841, 
				0.882)\\
				$r$ &  0.024 &
				0.02&
				(0.014, 
				0.052)&
				0.015 &
				0.013&
				(0.009, 
				0.033)\\
				\bottomrule
			\end{tabular}}
		\end{table}
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		\begin{figure}[p]
			\centering
			\includegraphics[width=.9\textwidth]{figure/zinbdensity-1} 
			\caption{Posterior density curves for each of the three variables. The flood posterior is shown in blue and the tornado posterior is shown in black.}
		\end{figure} 
	\end{frame}

>>>>

	\section{Zero Inflated Poisson Model}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}
		The zero inflated Poisson model has two parameters p, the probability of a structural zero, and $\lambda$, the usual Poisson parameter. 
		For a single X, the probability density is:
		$$pI_{X=0}(X) + (1-p)\frac{\lambda^Xe^-\lambda}{X!}$$
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}
		One way to write the likelihood is as follows:
		$$\prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]$$
		Bayarri, Berger, and Datta (2008)\footnotemark suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior:
		$$\prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i - 1/2}}{x_i!}\bigg]$$
		\footnotetext[1]{Bayarri, M., Berger, J., Datta, G. (2008). Objective testing of Poisson versus inflated Poisson models. IMS 	Collections, 3, 105-121.}
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}
		We can simplify a bit to get the following conditionals:
		\[
p(\lambda|X, p) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[e^{-\lambda}\lambda^{x_i - 1/2}\bigg]
\]

\[
p(p|X, \lambda) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\bigg]
\]

These are unrecognizable. I ended up alternatively sampling from the two using Metropolis Hastings. For lambda I settled on a gamma proposal, and for p I settled on a Beta proposal. 
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}
		\begin{table}
    \centering
    \caption{After taking 20,000 samples from each variable and discarding 10,000 as a burn-in, we obtained the following (values are the mean of the posterior sample followed by 95 percent credible interval in parentheses)}
    \label{t:rZIP}
    \begin{tabular}{| l | l | l |}
    \hline
    Event Type & $\lambda$ & p  \\ \hline
    Tornado & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
    Flash Flood & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
    \end{tabular}
\end{table}
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}
		Convergence of parameters (flood model):
		\centering
		\includegraphics[width=.9\textwidth]{figure/ZIP_Flood_Conv.pdf} 
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}
		Posterior Densities (Blue = Flood; Black = Tornado):
		\centering
		\includegraphics[width=.9\textwidth]{figure/ZIP_Density.pdf} 
	\end{frame}	
	\section{Zero Inflated Negative Binomial Model}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		The Zero Inflated Negative Binomial (ZINB) model has three parameters $\sigma,$ the probability of a structural zero, and $p,r$ the usual negative binomial parameters. 
		For a single $X$ the probability density is: 
		$$p(X|\sigma, p, r) = \sigma I_{X=0}(X) + (1-\sigma)\frac{\Gamma(r+X)}{\Gamma(r)X!}p^X(1-p)^r$$
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		We took the uniform priors on $\sigma$ and $p$ and the non-informative prior $r^{-1/2}$ on $r$.
		This gave a posterior of:
		\scriptsize$$p(r,\sigma, p|X)\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}\prod_{i=1}^N\left(\frac{\Gamma(r+X_i)}{r^{1/2}\Gamma(r)}\right)$$
		\normalsize This factors into nothing. Therefore we used the Metropolis Hastings Algorithm with proposal distributions of truncated normals.
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model:Results}
		\begin{itemize}
			\item Curse of Dimensionality; very slow convergence.
			\item High autocorrelation
			\item Low confidence in results
		\end{itemize}
		\begin{table}
			\centering
			\scalebox{0.6}{
			\begin{tabular}{lcccccc}
				\toprule
				&\multicolumn{3}{c}{Flash Flood}&\multicolumn{3}{c}{Tornado}\\
				\cmidrule(r){2-4}\cmidrule(l){5-7}
				Parameter & Mean Value &Median Value & (95\% CI)& Mean Value &Median Value & (95\% CI)\\
				\midrule
				$\sigma$ & 0.307 &
				0.282&
				(0.013, 
				0.716)&
				0.299 &
				0.276&
				(0.012, 
				0.702)\\
				$p$ &  0.554 &
				0.555&
				(0.518, 
				0.589)&
				0.862 &
				0.863&
				(0.841, 
				0.882)\\
				$r$ &  0.024 &
				0.02&
				(0.014, 
				0.052)&
				0.015 &
				0.013&
				(0.009, 
				0.033)\\
				\bottomrule
			\end{tabular}}
		\end{table}
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		\begin{figure}[p]
			\centering
			\includegraphics[width=.9\textwidth]{figure/zinbdensity-1} 
			\caption{Posterior density curves for each of the three variables. The flood posterior is shown in blue and the tornado posterior is shown in black.}
		\end{figure} 
	\end{frame}
